# Good Workflow practices: Guerrilla analytics

Within all work there are good practices and rules to follow. Which can also be called workflows. Datascience is no different, and a book  written about the matter is [Guerrilla Analytics](https://guerrilla-analytics.net/). in here Enda Ridge talks about 7 principles to uphold while working on a datascience project. These principle help keep the project understandable, explainable and reproducible. 

#### the Guerrila analytics:
 >1. Space is cheap, confusion is expensive
 >1. Prefer simple, visual project structures and conventions
 >3. Prefer automation with program code
 >4. Maintain a link between data on the file system, data in the analytics environment, and data in work products
 >5. Version control changes to data and analytics code
 >6. Consolidate team knowledge in version-controlled builds  
 >7. Prefer analytics code that runs from start to finish
 
 To help me uphold the first 3 principles i automated the creation of a new project folder with a simple R script:
 
```{r, eval=FALSE}
#create a new project in current directory
library(usethis) #2.0.1
currect_dir <- getwd()
go <- readline(prompt = paste("Do you want to creat a new project folder at", currect_dir, "? y/n: ", sep = " "))

if (go == "y"){

  #make project folder
  project.name <- readline(prompt = "Enter project name: ")
  dir.create(project.name)
  for (i in c("data", "doc", "external", "output", "src")){
    dir.create(paste0(project.name, "/", i))
  }

  #make a project
  usethis::create_project(project.name)

  # #connect to git
  # usethis::use_git()
  # usethis::use_github()


  } else {
    stop()}
```
 
 this script creates a standard file structure:

```{r}
dir_tree("/Users/larsd/Documents/R/Project/", recurse = TRUE)
```
 
 This structure will also be applied to the project folder that contains the files for this bookdown portfolio

```{r}
dir_tree(recurse = TRUE)
```





