# Portfolio assignment 7

## 1. Load the flu (“./data/flu_data.csv), the dengue (.”/data/dengue_data.csv) and the gapminder ({dslabs} package) into three separate dataframes in R

```{r}
dengue <- read.csv(here::here("7_data/dengue_data.txt"), skip = 11) 
head(dengue)

flu <- read.csv(here::here("7_data/flu_data.txt"), skip = 11)
head(flu)

head(gapminder)
```

## 2. Check if they are in the right shape. Is the data in the ‘tidy’ format? If not change the format to ‘tidy’

```{r}
dengue <- dengue %>% pivot_longer(Argentina:Venezuela, names_to = "country", values_to = "value")
head(dengue)

flu <- flu %>% pivot_longer(Argentina:Uruguay, names_to = "country", values_to = "value")
head(flu)

```


## 3. Change the country and date variables of the three tables so that they coincide in terms of data type, class and values

```{r}
dengue <- dengue %>% separate("Date", into = c("year", "month", "day"), sep = "-") 

dengue <- dengue[-c(2,3)] 
dengue$year <- as.integer(dengue$year)
dengue$country <- as.factor(dengue$country)

dengue <- dengue %>% 
            group_by(year) %>% 
            summarise("country" = country, "dengue activity" = sum(value, na.rm = T))
head(dengue)


flu <- flu %>% separate("Date", into = c("year", "month", "day"), sep = "-") 

flu <- flu[-c(2,3)] 
flu$year <- as.integer(flu$year)
flu$country <- as.factor(flu$country)

flu <- flu %>% 
        group_by(year) %>% 
        summarise("country" = country, "influenza activity" = sum(value, na.rm = T))

flu
```

## 4. Store the three tables as separate (so six in total) .csv and .rds files.

```{r}
for (x in c("flu", "dengue", "gapminder")) {
  filecsv <- paste0(paste0(here("7_data", x)), "_tidy.csv") 
  filerds <- paste0(paste0(here("7_data", x)), "_tidy.rds")
  write.csv(get(x), file = filecsv)
  saveRDS(get(x), file = filerds)
}

```


## 5. In Dbeaver create a new PostgreSQL database “workflowsdb”
### 6. Using RPostgreSQL, insert the tables into the database.

```{r, eval=FALSE}
con <- dbConnect(RPostgres::Postgres(), 
                 dbname = "workflowsdb", 
                 host="localhost", 
                 port="5432", 
                 user="postgres", 
                 password="NEWkidsweek123") 

dbWriteTable(con, "dengue data", dengue)
dbWriteTable(con, "flu data", flu)
dbWriteTable(con, "gapminder data", gapminder)

dbDisconnect(con)
```


Inspect the contents of the tables with SQL (in DBeaver) and save the SQL script.

```{sql connection= con}
SELECT count(* | *),  FROM "dengue data";

```

```{sql connection= con}
SELECT SUM("dengue activity") FROM "dengue data";

```
```{sql connection=con}
SELECT LENGTH("dengue activity") FROM "dengue data";
```




Inspect the contents of the tables with dplyr (in R) and save a RMarkdown showing what you are doing.

Load the gapminder data in R and change the dataframe in such as way that you could join it to dengue and flue.

Save this clean gapminder data in the “workflowsdb” database

Perform some joins (your choice) with SQL (can be done in DBeaver or with dplyr.

Generate a joined table, and export this from the database to R.

Show some descriptive statistics with this table, and at least 3 visualisations using ggplot2.

show all of your actions in this assignment in a Rmd file, perhaps with pictures and provide text explaining and showcasing your skills.

